Austral Spring Extratropical Southern Hemisphere zonally assymetric circulation \thanks{The research was supported by UBACyT20020170100428BA and the CLIMAX Project funded by Belmont Forum/ANR-15-JCL/-0002-01. Elio Campitelli was supported by a PhD grant from CONICET, Argentina.}

### 2021-08-17 analysis/paper/shceof.Rmd
Render

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  cache = TRUE,
  comment = "#>",
  cache.extra = 45,
  cache.path = "../cache/paper/",
  fig.path = "../figures/",
  dpi = 300,
  dev = c("pdf", "png")
)

>>>>

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  cache = TRUE,
  comment = "#>",
  cache.extra = 42,
  cache.path = "../cache/paper/",
  fig.path = "../figures/",
  dpi = 300,
  dev = c("pdf", "png")
)

### 2021-08-17 analysis/paper/shceof.tex
Render

One of the main features of the zonally asymmetric circulation are the Pacific--South American Patterns \citep{mo2001}, which are associated with precipitation anomalies in South America.
Besides, \citet{raphael2007} suggests that variability in the planetary wave 3 projected onto its climatological location is associated with Antarctic sea ice concentration anomalies.
\citet{irving2015} showed that planetary wave amplitude in general is associated sea ice concentration, as well as temperature and precipitation anomalies.

@@

\textcolor{red}{Los colores de la Tabla 1 no dicen mucho en la medida que no se conoce cuál sería la paleta. Quizás se pueden remarcar en negrita ambos aclarando que son significativos (lo son?)}.

@@

\caption{\label{tab:corr-ceof-splitted}Coefficient of determination (\(R^2\)) of the absolute magnitude of complex EOFs between 200 hPa and 50 hPa computing EOF separatedly for each level. Background colour indicates the value of \(R^2\).}
\centering
\begin{tabular}[t]{l>{}r>{}r>{}r}
\toprule
\multicolumn{1}{c}{} & \multicolumn{3}{c}{50 hPa} \\
\cmidrule(l{3pt}r{3pt}){2-4}
200 hPa & cEOF1 & cEOF2 & cEOF3\\
\midrule
cEOF1 & \cellcolor[HTML]{E1C0BB}{\textcolor{black}{0.28}} & \cellcolor[HTML]{FDFCFC}{\textcolor{black}{0.01}} & \cellcolor[HTML]{FDFAFA}{\textcolor{black}{0.02}}\\
cEOF2 & \cellcolor[HTML]{FFFFFF}{\textcolor{black}{0.00}} & \cellcolor[HTML]{BB7A73}{\textcolor{white}{0.60}} & \cellcolor[HTML]{FDFAFA}{\textcolor{black}{0.02}}\\
cEOF3 & \cellcolor[HTML]{FFFFFF}{\textcolor{black}{0.00}} & \cellcolor[HTML]{FFFFFF}{\textcolor{black}{0.00}} & \cellcolor[HTML]{FDFAFA}{\textcolor{black}{0.02}}\\
\bottomrule
\end{tabular}
\end{table}

@@

\textcolor{red}{se me vino una duda, ¿se pueden comparar de algún modo las amplitudes que mostras en esta figura? en caso de que si, creo que habría que mencionar algo sobre lo que pasa con el cEOF2 en la estratósfera} All cEOFs have non zero mean (Fig. \ref{fig:extended-series}) because the geopotential fields that enter into the cEOFs algorithm are anomalies with respect to the zonal mean instead of the time mean.
However, cEOF2's mean is almost zero, which indicates that only cEOF1 includes variability that significantly projects onto the mean zonally anomalous field.
There are no significant simultaneous correlation between the time series.
Both cEOFs show year-to-year variability but show no evidence of decadal variability.

\textcolor{red}{En la figura 3 agregaría dos cosas: i)el eje y en la columna derecha (si la queres ver más grande perdes la referencia) ii) creo que serviría agregar una líneas verticales en la grilla para poder tener una idea un poco mejor de a qué años corresponde más o menos cada valor. pueden ser un poco más finas que las que ya están, pero creo que sería una referencia util para visualmente poder decir a qué año corresponde cada punto. Además, fijate que la figura te quedo de un ancho mayor que el de la página y te quedo un "EOF" en el eje-y a la izquierda que no creo que sea lo más correcto.}

@@

\begin{figure}
\centering
\includegraphics{../figures/eof1-regr-gh-1.pdf}
\caption{\label{fig:eof1-regr-gh}Regression of SON geopotential height anomalies (meters) with the (column 1) real and (column 2) imaginary parts of the first cEOF for the 1979 -- 2019 period at (row a) 50 hPa and (row b) 200 hPa. These coefficients come from multiple linear regression involving the real and imaginary parts. Areas marked with dots have p-values smaller than 0.05 adjusted for False Detection Rate.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../figures/sam-eof-vertical-1.pdf}
\caption{\label{fig:sam-eof-vertical}Coefficient of determination between the real and imaginary part of each cEOF and the SAM, Asymmetric SAM (A-SAM) and Symmetric SAM (S-SAM) indices computed at each level according to \citet{campitelli2021}. Thick lines represent estimates with p-value \textless{} 0.01 corrected for False Detection Rate \citep{benjamini1995}.}
\end{figure}

@@

\caption{\label{tab:psa-eof2}Correlations between the Real and Imaginary parts of cEOF2 and the PSA1 and PSA2 modes computed as the second and third EOFs of seasonal geopotential height anomalies \citep[following][]{mo2001} for the 1979 -- 2019 period. 95\% confidence intervals in parenthesis. Background colour indicates the value of correlation.}
\centering
\begin{tabular}[t]{l>{}l>{}l}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{cEOF2} \\
\cmidrule(l{3pt}r{3pt}){2-3}
PC & Real & Imaginary\\
\midrule
PSA1 & \cellcolor[HTML]{C3BBDE}{\textcolor{black}{-0.32 (CI: -0.57 -- -0.015)}} & \cellcolor[HTML]{6259AA}{\textcolor{white}{-0.82 (CI: -0.9 -- -0.69)}}\\
PSA2 & \cellcolor[HTML]{9E4E47}{\textcolor{white}{0.81 (CI: 0.67 -- 0.9)}} & \cellcolor[HTML]{F3F2F8}{\textcolor{black}{-0.062 (CI: -0.36 -- 0.25)}}\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}
\centering
\includegraphics{../figures/o3-regr-1.pdf}
\caption{\label{fig:o3-regr}Regression of SON mean Total Ozone Column anomalies (shaded, Dobson Units) with the (a) real and (b) imaginary parts of the cEOF1 for the 1979 -- 2019 period. On contours, the mean zonal anomaly of Total Ozone Column (negative contours in dashed lines, Dobson Units). Areas marked with dots have p-values smaller than 0.05 adjusted for False Detection Rate.}
\end{figure}

@@

\begin{figure*}
\includegraphics{../figures/sst-psi-2-1} \caption{Regression maps of cEOF2 with SST (K, column a) and streamfunction zonal anomalies (\(m^2/s\times10^-7\), shaded) with their corresponding activity wave flux (vectors) (column b). Areas marked with dots have p-values smaller than 0.01 adjusted for FDR.}\label{fig:sst-psi-2}
\end{figure*}

@@

Since the Real cEOF2 represents just a different phase of the same wave train, it would be expected to show a similar forcing pattern to the Imaginary cEOF2 with a slight translation of its location.
However, Figure~\ref{fig:sst-psi-2}.a1 and b1 show that the Real cEOF2 is not associated either with any significant SST nor streamfunction anomalies in the tropics.
The correlation between the Real cEOF2 and ENSO is also not significant (0.00025 (CI: -0.31 -- 0.31)).
\textcolor{red}{Usaría notación científica por la cantidad de 0 que tiene}.
This lack of tropical signal suggest that the Real cEOF2 and the Imaginary cEOF2 have very different sources of variability.

@@

\textcolor{red}{Fig. 12 Revisa cómo te quedo abajo de la figura la indicación de qué son cada uno de los puntos (te quedaron dos circulos y una línea extra que no va)}

Furthermore, Figure~\ref{fig:enso-phase} suggest that highest amplitude cEOF2 years tend to coincide with strong ENSO years.
The correlation between the absolute magnitude of the ONI and the magnitude of the cEOF2 is 0.45 (CI: 0.17 -- 0.67).
This relationship, however, appears to be driven only by the three years with strongest ENSO events in the period (2015, 1997, and 1982) which also coincide with the three years with strongest cEOF2 magnitude.
If those years are removed, the correlation becomes non-significant (0.047 (CI: -0.28 -- 0.36)).
Furthermore, even when using all the datapoints, the Spearman correlation --which is robust to outliers-- is also non-significant (0.2, p-value= 0.21).
Therefore, the relationship between the magnitude of the cEOF2 train wave and ONI remains uncertain.

@@

\textcolor{red}{Revisa el caption de la figura 14 que equedo un estilo distinto que los anteriores. Te falto aclarar también qué es la caja (¿la inventaste vos o habías utilizado la de algún otro trabajo?). Fijate también de aumentar la separación entre la columna izquierda y derecha para que no se superpongan los labels. También fijate que esta barra de colores no te quedo con el borde externo como el resto. }

\begin{figure}
\centering
\includegraphics{../figures/pp-oceania-1.pdf}
\caption{\label{fig:pp-oceania}Regression of SON mean precipitation anomalies in Oceania (mm per day, shaded) and (column a) cEOF1 the (row 1) Imaginary and (column 1) Real phase. For the 1979 -- 2019. Areas marked with dots have p-values smaller than 0.05 adjusted for False Detection Rate. In black is the West Australia region, defined as the box between 130°E and 155°E and 40°S and 15°S.}
\end{figure}

The precipitation anomalies in Oceania associated with cEOF1 are weak and not statistically significant (Figure~\ref{fig:pp-oceania} column a) whereas cEOF2 is associated with large and widespread anomalies. The Real cEOF2 is associated with positive seasonal anomalies in Northern and Eastern Australia, and negative ones over New Zealand, while precipitation anomalies associated with Imaginary cEOF2 are negative over all Eastern Australia. As it could be expected, these negative anomalies are similar to the springtime precipitation anomalies observed in relation with the Asymmetric SAM \citep{campitelli2021} and ENSO \citep{cai2011}.
\textcolor{red}{Acá te faltaría  sacar un poco más el jugo asociando con la figura 11. Además, ¿revisaste si podes también usar algún resultado asociado a los PSA? }

\textcolor{red}{Revisa lo mismo que para la otra figura del tipo. }

@@

Figure~\ref{fig:pp-america} shows regression maps of seasonal precipitation with each cEOF in South America. cEOF1 is associated with negative regressions coefficients for precipitation in Southern Brazil and Paraguay, although these are not statistically significant (Fig.~\ref{fig:pp-america} column a). The strongest precipitation regression coefccients are the ones associated with the Imaginary cEOF2. The positive regression coefficients on Southeastern South America (SESA) and Central Chile, and negative ones over South Atlantic Convergence Zone is a well-known springtime precipitation signature of ENSO \citep{cai2020a} and it is also similar to the precipitation anomalies associated with the A-SAM \citep{campitelli2021}.
\textcolor{red}{Lo mismo que te puse antes. Te faltaría  sacar un poco más el jugo asociando con la figura 11. Además, ¿revisaste si podes también usar algún resultado asociado a los PSA? }

>>>>

One of the main features of the zonally asymmetric circulation are the Pacific--South American Patterns \citep{mo2001}, which are associated with precipitation anomalies in South America, particularly in the September-October-November trimester (cita).
Besides, \citet{raphael2007} suggests that variability in the planetary wave 3 projected onto its climatological location is associated with Antarctic sea ice concentration anomalies.
\citet{irving2015} showed that planetary wave amplitude in general is associated sea ice concentration, as well as temperature and precipitation anomalies.

@@

@@

\caption{\label{tab:corr-ceof-splitted}Coefficient of determination (\(R^2\)) of the absolute magnitude of complex EOFs between 200 hPa and 50 hPa computing EOF separately for each level. p-values lower than 0.01 in bold.}
\centering
\begin{tabular}[t]{l>{}r>{}r>{}r}
\toprule
\multicolumn{1}{c}{} & \multicolumn{3}{c}{50 hPa} \\
\cmidrule(l{3pt}r{3pt}){2-4}
200 hPa & cEOF1 & cEOF2 & cEOF3\\
\midrule
cEOF1 & \textbf{0.28} & 0.01 & 0.02\\
cEOF2 & 0.00 & \textbf{0.60} & 0.02\\
cEOF3 & 0.00 & 0.00 & 0.02\\
\bottomrule
\end{tabular}
\end{table}

@@

All cEOFs have non zero mean (Fig. \ref{fig:extended-series}) because the geopotential fields that enter into the cEOFs algorithm are anomalies with respect to the zonal mean instead of the time mean.
However, cEOF2's mean is almost zero, which indicates that only cEOF1 includes variability that significantly projects onto the mean zonally anomalous field.
There are no significant simultaneous correlation between the time series.
Both cEOFs show year-to-year variability but show no evidence of decadal variability.

@@

\begin{figure}
\centering
\includegraphics{../figures/eof1-regr-gh-1.pdf}
\caption{\label{fig:eof1-regr-gh}Regression of SON geopotential height anomalies (\(m^2s^{-1}\)) with the (column 1) real and (column 2) imaginary parts of the first cEOF for the 1979 -- 2019 period at (row a) 50 hPa and (row b) 200 hPa. These coefficients come from multiple linear regression involving the real and imaginary parts. Areas marked with dots have p-values smaller than 0.01 adjusted for False Detection Rate.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../figures/sam-eof-vertical-1.pdf}
\caption{\label{fig:sam-eof-vertical}Coefficient of determination between the real and imaginary part of each cEOF and the SAM, Asymmetric SAM (A-SAM) and Symmetric SAM (S-SAM) indices computed at each level according to \citet{campitelli2021} for the 1979 -- 2019 period. Thick lines represent estimates with p-value \textless{} 0.01 corrected for False Detection Rate \citep{benjamini1995}.}
\end{figure}

@@

\caption{\label{tab:psa-eof2}Correlations between the Real and Imaginary parts of cEOF2 and the PSA1 and PSA2 modes computed as the second and third EOFs of seasonal geopotential height anomalies \citep[following][]{mo2001} for the 1979 -- 2019 period. 95\% confidence intervals in parenthesis. p-values lower than 0.01 in bold.}
\centering
\begin{tabular}[t]{l>{}l>{}l}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{cEOF2} \\
\cmidrule(l{3pt}r{3pt}){2-3}
PC & Real & Imaginary\\
\midrule
PSA1 & -0.32 (CI: -0.57 -- -0.02) & \textbf{-0.82 (CI: -0.9 -- -0.69)}\\
PSA2 & \textbf{0.81 (CI: 0.67 -- 0.9)} & -0.06 (CI: -0.36 -- 0.25)\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}
\centering
\includegraphics{../figures/o3-regr-1.pdf}
\caption{\label{fig:o3-regr}Regression of SON mean Total Ozone Column anomalies (shaded, Dobson Units) with the (a) real and (b) imaginary parts of the cEOF1 for the 1979 -- 2019 period. On contours, the mean zonal anomaly of Total Ozone Column (negative contours in dashed lines, Dobson Units). Areas marked with dots have p-values smaller than 0.01 adjusted for False Detection Rate.}
\end{figure}

@@

\begin{figure*}
\includegraphics{../figures/sst-psi-2-1} \caption{Regression maps of cEOF2 with SST (K, column a) and streamfunction zonal anomalies (\(m^2/s\times10^-7\), shaded) with their corresponding activity wave flux (vectors) (column b) for the 1979 -- 2019 period. Areas marked with dots have p-values smaller than 0.01 adjusted for FDR.}\label{fig:sst-psi-2}
\end{figure*}

@@

Since the Real cEOF2 represents just a different phase of the same wave train, it would be expected to show a similar forcing pattern to the Imaginary cEOF2 with a slight translation of its location.
However, Figure~\ref{fig:sst-psi-2}.a1 and b1 show that the Real cEOF2 is not associated either with any significant SST nor streamfunction anomalies in the tropics.
The correlation between the Real cEOF2 and ENSO is also not significant (0 (CI: -0.31 -- 0.31)).
This lack of tropical signal suggest that the Real cEOF2 and the Imaginary cEOF2 have very different sources of variability.

@@

Furthermore, Figure~\ref{fig:enso-phase} suggest that highest amplitude cEOF2 years tend to coincide with strong ENSO years.
The correlation between the absolute magnitude of the ONI and the magnitude of the cEOF2 is 0.45 (CI: 0.17 -- 0.67).
This relationship, however, appears to be driven only by the three years with strongest ENSO events in the period (2015, 1997, and 1982) which also coincide with the three years with strongest cEOF2 magnitude.
If those years are removed, the correlation becomes non-significant (0.05 (CI: -0.28 -- 0.36)).
Furthermore, even when using all the datapoints, the Spearman correlation --which is robust to outliers-- is also non-significant (0.2, p-value= 0.21).
Therefore, the relationship between the magnitude of the cEOF2 train wave and ONI remains uncertain.

@@

\begin{figure}
\centering
\includegraphics{../figures/pp-oceania-1.pdf}
\caption{\label{fig:pp-oceania}Regression of SON mean precipitation anomalies in Oceania (mm per day, shaded) with the (row a) real and (row b) imaginary pars of the (column 1) cEOF1 and (column 2) cEOF2. For the 1979 -- 2019. Areas marked with dots have p-values smaller than 0.01 adjusted for False Detection Rate. In black is the West Australia region, defined as the box between 130°E and 155°E and 40°S and 15°S.}
\end{figure}

The precipitation anomalies in Oceania associated with cEOF1 are weak and not statistically significant (Figure~\ref{fig:pp-oceania} column 1) whereas cEOF2 is associated with large and widespread anomalies. The Real cEOF2 is associated with positive seasonal anomalies in Northern and Eastern Australia, and negative ones over New Zealand, while precipitation anomalies associated with Imaginary cEOF2 are negative over all Eastern Australia. As it could be expected, these negative anomalies are similar to the springtime precipitation anomalies observed in relation with the Asymmetric SAM \citep{campitelli2021} and ENSO \citep{cai2011}.
\textcolor{red}{Acá te faltaría  sacar un poco más el jugo asociando con la figura 11. Además, ¿revisaste si podes también usar algún resultado asociado a los PSA? }

@@

Figure~\ref{fig:pp-america} shows regression maps of seasonal precipitation with each cEOF in South America. cEOF1 is associated with negative regressions coefficients for precipitation in Southern Brazil and Paraguay, although these are not statistically significant (Fig.~\ref{fig:pp-america} column 1). The strongest precipitation regression coefficients are the ones associated with the Imaginary cEOF2. The positive regression coefficients on Southeastern South America (SESA) and Central Chile, and negative ones over South Atlantic Convergence Zone is a well-known springtime precipitation signature of ENSO \citep{cai2020a} and it is also similar to the precipitation anomalies associated with the A-SAM \citep{campitelli2021}.
\textcolor{red}{Lo mismo que te puse antes. Te faltaría  sacar un poco más el jugo asociando con la figura 11. Además, ¿revisaste si podes también usar algún resultado asociado a los PSA? }
